WASP-4 is accelerating towards the Earth

### 2020-01-22 paper/ms.tex
paper minor

\documentclass[RNAAS]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}

@@

% NOTE: may want to improve title
\title{WASP-4 is accelerating towards the Earth}

@@

%% Note that RNAAS manuscripts DO NOT have abstracts.
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{Exoplanet tides (497), Exoplanet dynamics (490), Radial velocity 
(1332), Transit timing variation method (1710)}

%% Start the main body of the article. If no sections in the 
%% research note leave the \section call blank to make the title.

@@

Since the first discovery of transiting hot Jupiters, the hunt for
tidal orbital decay has received much attention (CITE LIKE 20 PAPERS).
These efforts are gradually beginning to yield fruit: by combining
transit and occultation timing and radial velocity measurements, CITET
Yee have shown that the data for the hot Jupiter WASP-12b are most
compatible with orbital decay.

This study highlights a point that, while obvious, has perhaps not
yet received due attention.
The point is that the hunt for orbital decay of hot Jupiters will be
crippled without long-term radial velocity monitoring programs.
The reason is simple: when the timescale of orbital period change far
exceeds the observing baseline, the first
deviation from constant periodicity is always quadratic in the transit
number:
\begin{align}
  t_{\rm tra} = t_0 + P E + \frac{1}{2}\frac{dP}{dE}E^2,
\end{align}
where
\begin{align}
  \frac{dP}{dE} = P \frac{dP}{dt},
\end{align}
$E$ is the transit number, $P$ is the orbital period, $t_0$ is the
reference epoch, and $t_{\rm tra}$ is the transt mid-time.

@@

The main example of this study is the hot Jupiter WASP-4b which has an
orbital period that from transits appears to be decreasing by about 10
milliseconds per year.
After discovering the timing variations in the TESS data, we obtained
N additional radial velocity measurements using Keck-HIRES,
extending the RV baseline by Y years.
Previously, the radial velocities were sparse and were consistent with
not showing any linear trend.
Our new measurements reveal a line of sight acceleration of
$\dot{\gamma} = -XX {\rm m\,s^{-1}\,yr^{-1}}$.
This translates to an expected period decrease simply from the light
travel time effect of ZZ milliseconds per year --- about what is
observed.

\section{WASP-4's acceleration towards the Earth}

>>>>

\documentclass[12pt,twocolumn,tighten]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{\today}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{WASP-4 is accelerating towards the Earth}

@@

% TODO: may want to improve title
\title{WASP-4 is accelerating towards the Earth}

@@

\begin{abstract}
  We recently noted that combining measurements from the Transiting
  Exoplanet Survey Satellite (TESS) with ground-based transit
  observations, the orbital period of WASP-4b seemed to be decreasing.
  Possible explanations included tidal orbital decay, orbital
  precession, and light-travel time effects.
  In this work we present new observations of the radial velocity of
  WASP-4, acquired with Keck-HIRES.
  They show that the system is accelerating towards our line of sight,
  at $\dot{\gamma} = X.XX \pm Y.YY\ {\rm m}\,{\rm s}^{-1}\,{\rm
  yr}^{-1}$.
  This implies a period decrease of X.XX milliseconds per year,
  compared to the observed value of Y.YY.
  Combining the radial velocities with new speckle imaging from
  Gemini-Zorro, we show that the acceleration is best explained by a
  10-200$\,M_{\rm Jup}$ companion between 10 and 100$\,$AU.
  We encourage further long-term monitoring of hot Jupiter radial
  velocities; these observations are critical in the ongoing hunt
  for tidal orbital decay.
\end{abstract}

\keywords{Exoplanet tides (497), Exoplanet dynamics (490), Radial velocity 
(1332), Transit timing variation method (1710)}

@@

Since the first discovery of transiting hot Jupiters, the hunt for
tidal orbital decay has received much attention (CITE LIKE 20 PAPERS).
These efforts are gradually beginning to yield fruit: by combining
transit and occultation timing and radial velocity measurements, CITE
Yee et al{.} have shown that the data for the hot Jupiter WASP-12b are
most compatible with orbital decay.

The present study highlights a point that, while obvious, has perhaps
not yet received due attention.  The point is that the hunt for
orbital decay of hot Jupiters will be crippled without long-term
radial velocity monitoring programs.
The reason is simple: when the timescale of orbital period change far
exceeds the observing baseline, the first deviation from constant
periodicity is always quadratic in the transit number:
\begin{align}
  t_{\rm tra} = t_0 + P E + \frac{1}{2}\frac{dP}{dE}E^2,
\end{align}
where
\begin{align}
  \frac{dP}{dE} = P \frac{dP}{dt},
\end{align}
$E$ is the transit number, $P$ is the orbital period, $t_0$ is the
reference epoch, and $t_{\rm tra}$ is the transt mid-time.

@@

The main focus of this study is the hot Jupiter WASP-4b, which has an
orbital period that from transits appears to be decreasing by about 10
milliseconds per year.
After discovering the timing variations in the TESS data (CITE Bouma
2019), we obtained N additional radial velocity measurements using
Keck-HIRES, extending the RV baseline by Y years.
Previously, the radial velocities were sparse and were consistent with
not showing any linear trend.
Our new measurements reveal a line of sight acceleration of
$\dot{\gamma} = -XX {\rm m\,s^{-1}\,yr^{-1}}$.
This translates to an expected period decrease simply from the light
travel time effect of ZZ milliseconds per year --- about what is
observed.

% FIXME: figure out correct sections
\S~\ref{sec:observations} presents the new observations...
\S~\ref{sec:timing} describes our timing analysis...
\S~\ref{sec:implications} ...
\S~\ref{sec:future}...

\section{WASP-4's acceleration towards the Earth}
\label{sec:observations}
